---
title: "Create household couples"
author: "Jenny"
date: "2020-01-28"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: console
---

### Household couples were created as follows:

```{r setup, include=FALSE}

options(scipen=999)
source(here::here("code/packages.R"))
source(here::here("code/settings.R"))
# load drake objects
loadd(hh_pairs,
      household_info,
      hh_pairs_filter,
      hh_pairs_kin,
      household_relationships
)

pairs <- hh_pairs_kin 
field <- relatedness_field 

pairs$sex1 <- household_relationships[["sex"]][match(pairs[["HOUSEHOLD_MEMBER1"]], household_relationships$userId)]
pairs$sex2 <- household_relationships[["sex"]][match(pairs[["HOUSEHOLD_MEMBER2"]], household_relationships$userId)]

#merge(pairs, household_relationships, by.x=, by.y=)
pairs_full_sex <- pairs[which(!is.na(pairs$sex1) & !is.na(pairs$sex2)),]

pairs_unrelated <- subset(pairs_full_sex, pairs_full_sex$kinship<0.05)
pairs_unrelated_hetero <- pairs_unrelated[-which(pairs_unrelated$sex1==pairs_unrelated$sex2),]
pairs_unrelated_hetero2 <- pairs_unrelated_hetero[-which(is.na(pairs_unrelated_hetero$sex1) | is.na(pairs_unrelated_hetero$sex2)),]

temp1 <- pairs_unrelated_hetero[which(pairs_unrelated_hetero$sex1==0),c(2,1,3,4,6,5)]
colnames(temp1) <- colnames(pairs_unrelated_hetero)

temp2 <- pairs_unrelated_hetero[which(pairs_unrelated_hetero$sex1==1),]
pairs_filter <- rbind(temp2, temp1)
  ## 79150      6

pairs_filter$house_rel_member1 <- household_relationships[[field]][match(pairs_filter[["HOUSEHOLD_MEMBER1"]], household_relationships$userId)]
pairs_filter$house_rel_member2 <- household_relationships[[field]][match(pairs_filter[["HOUSEHOLD_MEMBER2"]], household_relationships$userId)]
pairs_filter2 <- pairs_unrelated_hetero[which(pairs_filter$house_rel_member1==TRUE & pairs_filter$house_rel_member2==TRUE ) ,]
  ## 75048     6

colnames(pairs_filter2)[5] <- "HOUSEHOLD_MEMBER1_sex"
colnames(pairs_filter2)[6] <- "HOUSEHOLD_MEMBER2_sex"


init_pairs <- (dim(hh_pairs)[1])/2
sex_filter <- (dim(pairs_full_sex)[1])
relatedness_filter <- dim(pairs_unrelated)[1]
hetero_filter <- dim(pairs_unrelated_hetero)[1]
questionaire_filter <- dim(pairs_filter2)[1]
```


1. Household pairs were created using the file located here: `r {paste0(UKBB_dir,"/pheno/ukb6881.csv")}` and extracting only households where *exactly* two participants belonged to the same household. 
This resulted in `r dim(hh_pairs)[1]` individuals, representing `r init_pairs` household pairs. 

2. Sex was assigned to each member and pairs where either participant had sex information were removed: `r sex_filter`. 

3. Kinship was assigned to each pair using the kinship file (`r paste0(UKBB_dir,"/geno/","ukb1638_rel_s488366.dat")`). Pairs with kinship < 0.05 were retained, resulting in `r relatedness_filter` pairs. 

4. Pairs of the same sex were removed: 

5. Using the relatedness file: `r relations_file`, and corresponding relatedness field (`r relatedness_field`) corresponding to the question: "...." 



```{r filter}
head(hh_pairs_filter)

head(hh_pairs_kin)
```
